"Reproducible Research: Peer Assessment 1"

## Setting up working directory
```{r}
getwd()
setwd(dir ="D:/DATA ANALYSTS/R/R Course Directory")
dir()
```

## Install necessary packages for loading, pre-processing, and transforming the data.
```{r setup, include=FALSE}
options(repos = c(CRAN = "https://cran.rstudio.com/"))
install.packages("tidyverse")
library(tidyverse)
```

## Loading and preprocessing the data
```{r loaddata}
unzip(zipfile="activity.zip")
data <- read.csv("activity.csv")
```

## What is mean total number of steps taken per day?
```{r}
library(ggplot2)
total.steps <- tapply(data$steps, data$date, FUN=sum, na.rm=TRUE)
qplot(total.steps, binwidth=1000, xlab="total number of steps taken each day")
mean(total.steps, na.rm=TRUE)
median(total.steps, na.rm=TRUE)
```

## What is the average daily activity pattern?
```{r}
library(ggplot2)
averages <- aggregate(x=list(steps=data$steps), by=list(interval=data$interval),
                      FUN=mean, na.rm=TRUE)
ggplot(data=averages, aes(x=interval, y=steps)) +
    geom_line() +
    xlab("5-minute interval") +
    ylab("average number of steps taken")
```

## Imputing missing values
```{r how_many_missing}
missing <- is.na(data$steps)
# How many missing
table(missing)
```

## Are there differences in activity patterns between weekdays and weekends?
```{r}
weekday.or.weekend <- function(date) {
  day <- weekdays(date)
  if (day %in% c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday"))
    return("weekday")
  else if (day %in% c("Saturday", "Sunday"))
    return("weekend")
  else
    stop("invalid date")
}
filled.data$date <- as.Date(filled.data$date)
filled.data <- filled.data %>%
  mutate(day_type = ifelse(weekdays(date) %in% c("Saturday", "Sunday"), "weekend", "weekday"))
```

# Load necessary packages
```{r}
if (!requireNamespace("dplyr", quietly = TRUE)) {
  install.packages("dplyr")
}
library(dplyr)
```

```{r}
avg_steps <- filled.data %>%
  group_by(interval, day_type) %>%
  summarize(steps = mean(steps, na.rm = TRUE))
```

# Load ggplot2 if not already loaded
```{r}
if (!requireNamespace("ggplot2", quietly = TRUE)) {
  install.packages("ggplot2")
}
library(ggplot2)
```

# Create a panel plot
```{r}
ggplot(avg_steps, aes(x = interval, y = steps)) +
  geom_line(aes(color = day_type), type = "l", size = 1) +
  facet_grid(day_type ~ ., scales = "free_y") +
  xlab("5-minute interval") +
  ylab("Average number of steps taken") +
  ggtitle("Time Series Plot of Average Steps per 5-Minute Interval") +
  theme_minimal()
```
